---
title: "Statistics-PCA"
author: "Angela Zoss"
date: "July 23, 2016"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    fig_width: 10
    code_folding: show
    keep_md: true
---
# Multiple Factor Analysis

## Setting up the environment.

```{r, message=FALSE}
#setwd("/Users/az49/Box Sync/PhD-AngelaZoss/1-Opinion Survey/data/OpinSurv-DataAnalysis")

require(FactoMineR)
require(missMDA)
require(tidyverse)
require(gridExtra)
require(grid)
#require(data.table)
```

## First try: load data in string version and factor it. Could try to run the MFA directly.  Works okay, but doesn't deal with skipped responses, which requires missMDA and a different data format.

```{r}

#pca_table.csv already reduced to 51 observations
DF.PCA <- read.csv("pca_table.csv", header=TRUE, stringsAsFactors=FALSE, row.names=1)

DF.PCA$Q15.Demo.HighestDegree <- sub("’","", DF.PCA$Q15.Demo.HighestDegree)

DF.PCA$Q5.Demo.PrimaryAcademicField <- sub("-99","Skipped", DF.PCA$Q5.Demo.PrimaryAcademicField)

DF.PCA$Q5.Demo.PrimaryAcademicField <- factor(DF.PCA$Q5.Demo.PrimaryAcademicField)

DF.PCA$Q15.Demo.HighestDegree <- factor(DF.PCA$Q15.Demo.HighestDegree, 
                                    levels=c("Bachelors degree",
                                             "Masters degree",
                                             "Professional degree", 
                                             "Doctorate degree", 
                                             "Other"),
                                    labels=c("Bachelor's",
                                             "Master's",
                                             "Professional",
                                             "Doctoral",
                                             "Other"),
                                    ordered = TRUE)

freq4 <- c("None","A little", "Some", "A lot")

DF.PCA$Q1.Demo.NetworkConsumerExperience <-
  factor(DF.PCA$Q1.Demo.NetworkConsumerExperience, 
         levels=freq4,
         ordered = TRUE)

DF.PCA$Q21.Demo.NetworkProducerExperience <-
  factor(DF.PCA$Q21.Demo.NetworkProducerExperience, 
         levels=freq4,
         ordered = TRUE)

agree4 <- c("Not well at all","Not very well", "Somewhat well", "Very well")

DF.PCA$Q3.Demo.NetworkAnalysisResearch <-
  factor(DF.PCA$Q3.Demo.NetworkAnalysisResearch, 
         levels=agree4,
         labels=c("Not well\nat all","Not very\nwell", "Somewhat\nwell", "Very well"),
         ordered = TRUE)

DF.PCA$Q4.Demo.NetworkVisualizationResearch <-
  factor(DF.PCA$Q4.Demo.NetworkVisualizationResearch, 
         levels=agree4,
         labels=c("Not well\nat all","Not very\nwell", "Somewhat\nwell", "Very well"),
         ordered = TRUE)

freq5 <- c("Skipped","Never","Rarely", "Sometimes", "Most of the time", "Always")

DF.PCA$Q11.Demo.HowOftenProduceVis <- 
  factor(DF.PCA$Q11.Demo.HowOftenProduceVis,
                    levels=freq5,
                    ordered = TRUE)

for (col in c(8:43)){
  DF.PCA[,col] <- factor(DF.PCA[,col],levels=c("Skipped","Unfamiliar","Never/almost never","Seldom/rarely","Often","Almost always/always"),ordered=TRUE)
}

for (col in c(44:77)){
  DF.PCA[,col] <- factor(DF.PCA[,col],levels=c("Skipped","Unfamiliar","Very low","Somewhat low","Somewhat high","Very high"),ordered=TRUE)
}

for (col in c(78:93)){
  DF.PCA[,col] <- factor(DF.PCA[,col],levels=c("Skipped","Unfamiliar","Never/almost never","Seldom/rarely","Often","Almost always/always"),ordered=TRUE)
}



```

## Prepare data for imputing instead.  Load string versions of data, but then replace with numbers and refactor with new labels.

```{r}
#pca_table.csv already reduced to 51 observations
DF.PCA.NA <- read.csv("pca_table.csv", header=TRUE, stringsAsFactors=FALSE, row.names=1)

DF.PCA.NA$Q5.Demo.PrimaryAcademicField <-
    ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "-99", NA, 
           ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Anthropology", 1, 
                  ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Business", 2, 
                         ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Communication studies", 3, 
                                ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Computer sciences", 4, 
                                       ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Economics", 5, 
                                              ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Engineering", 6, 
                                                     ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Geography", 7, 
                                                            ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "History", 8, 
                                                                   ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Information science", 9, 
                                                                          ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Mathematics", 10, 
                                                                                 ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Medicine", 11, 
                                                                                        ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Other", 12, 
                                                                                               ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Physics", 13, 
                                                                                                      ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Political science", 14, 
                                                                                                             ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Psychology", 15, 
                                                                                                                    ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Public administration", 16, 
                                                                                                                           ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField == "Sociology", 17, "Error"))))))))))))))))))
  

DF.PCA.NA$Q5.Demo.PrimaryAcademicField <- factor(DF.PCA.NA$Q5.Demo.PrimaryAcademicField,
                                           levels=c(1:17),
                                           labels=c("Anthropology", "Business", "Communication studies", "Computer sciences", "Economics", "Engineering", "Geography", "History", "Information science", "Mathematics", "Medicine", "Other", "Physics", "Political science", "Psychology", "Public administration", "Sociology"))

DF.PCA.NA$Q15.Demo.HighestDegree <- sub("’","", DF.PCA.NA$Q15.Demo.HighestDegree)

DF.PCA.NA$Q15.Demo.HighestDegree <- 
    ifelse(DF.PCA.NA$Q15.Demo.HighestDegree == "Other", 5, 
           ifelse(DF.PCA.NA$Q15.Demo.HighestDegree == "Skipped", NA, 
                  ifelse(DF.PCA.NA$Q15.Demo.HighestDegree == "Bachelors degree", 1, 
                         ifelse(DF.PCA.NA$Q15.Demo.HighestDegree == "Masters degree", 2, 
                                ifelse(DF.PCA.NA$Q15.Demo.HighestDegree == "Professional degree", 3, 
                                       ifelse(DF.PCA.NA$Q15.Demo.HighestDegree == "Doctorate degree", 4, "Error"))))))

#DF.PCA.NA$Q15.Demo.HighestDegree <- factor(DF.PCA.NA$Q15.Demo.HighestDegree)

DF.PCA.NA$Q15.Demo.HighestDegree <- factor(DF.PCA.NA$Q15.Demo.HighestDegree, 
                                    levels=c(1, 2, 3, 4, 5),
                                    labels=c("Bachelor's",
                                             "Master's",
                                             "Professional",
                                             "Doctoral",
                                             "Other"),
                                    ordered = TRUE)

DF.PCA.NA$Q1.Demo.NetworkConsumerExperience <-
  ifelse(DF.PCA.NA$Q1.Demo.NetworkConsumerExperience == "Unfamiliar", 0, 
           ifelse(DF.PCA.NA$Q1.Demo.NetworkConsumerExperience == "Skipped", NA, 
                  ifelse(DF.PCA.NA$Q1.Demo.NetworkConsumerExperience == "None", 1, 
                         ifelse(DF.PCA.NA$Q1.Demo.NetworkConsumerExperience == "A little", 2, 
                                ifelse(DF.PCA.NA$Q1.Demo.NetworkConsumerExperience == "Some", 3, 
                                       ifelse(DF.PCA.NA$Q1.Demo.NetworkConsumerExperience == "A lot", 4, "Error"))))))

freq4 <- c("None","A little", "Some", "A lot")

DF.PCA.NA$Q1.Demo.NetworkConsumerExperience <-
  factor(DF.PCA.NA$Q1.Demo.NetworkConsumerExperience, 
         levels=c(1,2,3,4),
         labels=freq4,
         ordered = TRUE)

DF.PCA.NA$Q21.Demo.NetworkProducerExperience <-
  ifelse(DF.PCA.NA$Q21.Demo.NetworkProducerExperience == "Unfamiliar", 0, 
           ifelse(DF.PCA.NA$Q21.Demo.NetworkProducerExperience == "Skipped", NA, 
                  ifelse(DF.PCA.NA$Q21.Demo.NetworkProducerExperience == "None", 1, 
                         ifelse(DF.PCA.NA$Q21.Demo.NetworkProducerExperience == "A little", 2, 
                                ifelse(DF.PCA.NA$Q21.Demo.NetworkProducerExperience == "Some", 3, 
                                       ifelse(DF.PCA.NA$Q21.Demo.NetworkProducerExperience == "A lot", 4, "Error"))))))

DF.PCA.NA$Q21.Demo.NetworkProducerExperience <-
  factor(DF.PCA.NA$Q21.Demo.NetworkProducerExperience, 
         levels=c(1,2,3,4),
         labels=freq4,
         ordered = TRUE)

DF.PCA.NA$Q3.Demo.NetworkAnalysisResearch <-
  ifelse(DF.PCA.NA$Q3.Demo.NetworkAnalysisResearch == "Unfamiliar", 0, 
           ifelse(DF.PCA.NA$Q3.Demo.NetworkAnalysisResearch == "Skipped", NA, 
                  ifelse(DF.PCA.NA$Q3.Demo.NetworkAnalysisResearch == "Not well at all", 1, 
                         ifelse(DF.PCA.NA$Q3.Demo.NetworkAnalysisResearch == "Not very well", 2, 
                                ifelse(DF.PCA.NA$Q3.Demo.NetworkAnalysisResearch == "Somewhat well", 3, 
                                       ifelse(DF.PCA.NA$Q3.Demo.NetworkAnalysisResearch == "Very well", 4, "Error"))))))

agree4 <- c("Not well at all","Not very well", "Somewhat well", "Very well")

DF.PCA.NA$Q3.Demo.NetworkAnalysisResearch <-
  factor(DF.PCA.NA$Q3.Demo.NetworkAnalysisResearch, 
         levels=c(1,2,3,4),
         labels=agree4,
         ordered = TRUE)

DF.PCA.NA$Q4.Demo.NetworkVisualizationResearch <-
  ifelse(DF.PCA.NA$Q4.Demo.NetworkVisualizationResearch == "Unfamiliar", 0, 
           ifelse(DF.PCA.NA$Q4.Demo.NetworkVisualizationResearch == "Skipped", NA, 
                  ifelse(DF.PCA.NA$Q4.Demo.NetworkVisualizationResearch == "Not well at all", 1, 
                         ifelse(DF.PCA.NA$Q4.Demo.NetworkVisualizationResearch == "Not very well", 2, 
                                ifelse(DF.PCA.NA$Q4.Demo.NetworkVisualizationResearch == "Somewhat well", 3, 
                                       ifelse(DF.PCA.NA$Q4.Demo.NetworkVisualizationResearch == "Very well", 4, "Error"))))))

DF.PCA.NA$Q4.Demo.NetworkVisualizationResearch <-
  factor(DF.PCA.NA$Q4.Demo.NetworkVisualizationResearch, 
         levels=c(1,2,3,4),
         labels=agree4,
         ordered = TRUE)

DF.PCA.NA$Q11.Demo.HowOftenProduceVis <-
  ifelse(DF.PCA.NA$Q11.Demo.HowOftenProduceVis == "Unfamiliar", 0, 
           ifelse(DF.PCA.NA$Q11.Demo.HowOftenProduceVis == "Skipped", NA, 
                  ifelse(DF.PCA.NA$Q11.Demo.HowOftenProduceVis == "Never", 1, 
                         ifelse(DF.PCA.NA$Q11.Demo.HowOftenProduceVis == "Rarely", 2, 
                                ifelse(DF.PCA.NA$Q11.Demo.HowOftenProduceVis == "Sometimes", 3, 
                                       ifelse(DF.PCA.NA$Q11.Demo.HowOftenProduceVis == "Most of the time", 4, 
                                              ifelse(DF.PCA.NA$Q11.Demo.HowOftenProduceVis == "Always", 5, "Error")))))))

freq5 <- c("Never","Rarely", "Sometimes", "Most of the time", "Always")

DF.PCA.NA$Q11.Demo.HowOftenProduceVis <- 
  factor(DF.PCA.NA$Q11.Demo.HowOftenProduceVis,
                    levels = c(1,2,3,4,5),
                    labels = freq5,
                    ordered = TRUE)

for (col in c(8:43)){
  DF.PCA.NA[,col] <- 
    ifelse(DF.PCA.NA[,col] == "Unfamiliar", 0, 
           ifelse(DF.PCA.NA[,col] == "Skipped", NA, 
                  ifelse(DF.PCA.NA[,col] == "Never/almost never", 1, 
                         ifelse(DF.PCA.NA[,col] == "Seldom/rarely", 2, 
                                ifelse(DF.PCA.NA[,col] == "Often", 3, 
                                       ifelse(DF.PCA.NA[,col] == "Almost always/always", 4, "Error"))))))
}

for (col in c(8:43)){
  DF.PCA.NA[,col] <- factor(DF.PCA.NA[,col],levels=c(0,1,2,3,4),labels=c("Unfamiliar","Never/almost never","Seldom/rarely","Often","Almost always/always"),ordered=TRUE)
}

for (col in c(44:77)){
  DF.PCA.NA[,col] <- 
    ifelse(DF.PCA.NA[,col] == "Unfamiliar", 0, 
           ifelse(DF.PCA.NA[,col] == "Skipped", NA, 
                  ifelse(DF.PCA.NA[,col] == "Very low", 1, 
                         ifelse(DF.PCA.NA[,col] == "Somewhat low", 2, 
                                ifelse(DF.PCA.NA[,col] == "Somewhat high", 3, 
                                       ifelse(DF.PCA.NA[,col] == "Very high", 4, "Error"))))))
}

for (col in c(44:77)){
  DF.PCA.NA[,col] <- factor(DF.PCA.NA[,col],levels=c(0,1,2,3,4),labels=c("Unfamiliar","Very low","Somewhat low","Somewhat high","Very high"),ordered=TRUE)
}

for (col in c(78:93)){
  DF.PCA.NA[,col] <- 
    ifelse(DF.PCA.NA[,col] == "Unfamiliar", 0, 
           ifelse(DF.PCA.NA[,col] == "Skipped", NA, 
                  ifelse(DF.PCA.NA[,col] == "Never/almost never", 1, 
                         ifelse(DF.PCA.NA[,col] == "Seldom/rarely", 2, 
                                ifelse(DF.PCA.NA[,col] == "Often", 3, 
                                       ifelse(DF.PCA.NA[,col] == "Almost always/always", 4, "Error"))))))
}

for (col in c(78:93)){
  DF.PCA.NA[,col] <- factor(DF.PCA.NA[,col],levels=c(0,1,2,3,4),labels=c("Unfamiliar","Never/almost never","Seldom/rarely","Often","Almost always/always"),ordered=TRUE)
}
```

## Test for people who skipped too many of the Importance/Estimability questions (redundant with refactored code, which should produce a pca_table file with only 51 participants, but have to run to keep column numbers correct later)
```{r}

DF.PCA.NA$TotalSkipped <- rowSums(is.na(DF.PCA.NA))

DF.PCA.NA$ActiveSkipped <- rowSums(is.na(DF.PCA.NA[44:77]))

# extra omit user "R_djvhaRoxinvZmw1"

DF.PCA.NA <- DF.PCA.NA[DF.PCA.NA$ActiveSkipped < 3,]

```

```{r}
# new variables

DF.PCA.NA$ActiveVHigh <- rowSums(DF.PCA.NA[,44:77] == "Very high", na.rm=TRUE)
DF.PCA.NA$ActiveSHigh <- rowSums(DF.PCA.NA[,44:77] == "Somewhat high", na.rm=TRUE)
DF.PCA.NA$ActiveSLow <- rowSums(DF.PCA.NA[,44:77] == "Somewhat low", na.rm=TRUE)
DF.PCA.NA$ActiveVLow <- rowSums(DF.PCA.NA[,44:77] == "Very low", na.rm=TRUE)
DF.PCA.NA$ActiveUnfam <- rowSums(DF.PCA.NA[,44:77] == "Unfamiliar", na.rm=TRUE)

#DF.PCA.NA$TotalActive <- rowSums(DF.PCA.NA[,c(95:100)], na.rm=TRUE)

DF.PCA.NA$AToolsUnfam <- rowSums(DF.PCA.NA[,8:25] == "Unfamiliar", na.rm=TRUE)

DF.PCA.NA$VToolsUnfam <- rowSums(DF.PCA.NA[,26:43] == "Unfamiliar", na.rm=TRUE)

DF.PCA.NA$LayoutsUnfam <- rowSums(DF.PCA.NA[,78:93] == "Unfamiliar", na.rm=TRUE)

DF.PCA.NA$has_phd <- ifelse(DF.PCA.NA$Q15.Demo.HighestDegree == "Doctoral",1,0)

DF.PCA.NA$has_phd <- factor(DF.PCA.NA$has_phd, labels=c("no Doctorate","Doctorate"))

Humanities <- c("Anthropology", "Arts", "Classics", "History", "Languages", "Literature", "Philosophy", "Religion")

SocialSciences <- c("Archaeology", "Communication studies", "Cultural and ethnic studies", "Economics", "Geography", "History", "Information science", "Linguistics", "Political science", "Psychology", "Sociology")

LifeSciences <- c("Biology", "Chemistry", "Earth sciences", "Physics", "Space sciences")

FormalSciences <- c("Mathematics", "Computer sciences", "Logic", "Statistics", "Systems science")

Professional <- c("Architecture and design", "Business", "Divinity", "Education", "Engineering", "Human physical performance and recreation", "Journalism, media studies and communication", "Law", "Library and museum studies", "Medicine", "Military sciences", "Public administration")

Other <- c("Other")

DF.PCA.NA$FieldGroups <- 
  ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField %in% Humanities, 1, 
         ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField %in% SocialSciences, 2,
                ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField %in% LifeSciences, 3,
                       ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField %in% FormalSciences, 4,
                              ifelse(DF.PCA.NA$Q5.Demo.PrimaryAcademicField %in% Professional, 5, 6)))))

DF.PCA.NA$FieldGroups <- factor(DF.PCA.NA$FieldGroups, 
                                levels=c(1:6),
                                labels=c("Humanities", "Social Sciences", "Life Sciences", "Formal Sciences", "Professional", "Other"))


```

## redundant test for participants who skipped Importance/Estimability questions, now on the "Skipped" dataset??
```{r}

DF.PCA$TotalSkipped <- rowSums(DF.PCA == "Skipped")

DF.PCA$ActiveSkipped <- rowSums(DF.PCA[44:77] == "Skipped")

# extra omit user "R_djvhaRoxinvZmw1"

DF.PCA <- DF.PCA[DF.PCA$ActiveSkipped < 3,]

```

```{r}

# new variables for Skipped dataset

DF.PCA$ActiveVHigh <- rowSums(DF.PCA[,44:77] == "Very high", na.rm=TRUE)
DF.PCA$ActiveSHigh <- rowSums(DF.PCA[,44:77] == "Somewhat high", na.rm=TRUE)
DF.PCA$ActiveSLow <- rowSums(DF.PCA[,44:77] == "Somewhat low", na.rm=TRUE)
DF.PCA$ActiveVLow <- rowSums(DF.PCA[,44:77] == "Very low", na.rm=TRUE)
DF.PCA$ActiveUnfam <- rowSums(DF.PCA[,44:77] == "Unfamiliar", na.rm=TRUE)

#DF.PCA$TotalActive <- rowSums(DF.PCA[,c(95:100)], na.rm=TRUE)

DF.PCA$AToolsUnfam <- rowSums(DF.PCA[,8:25] == "Unfamiliar", na.rm=TRUE)

DF.PCA$VToolsUnfam <- rowSums(DF.PCA[,26:43] == "Unfamiliar", na.rm=TRUE)

DF.PCA$LayoutsUnfam <- rowSums(DF.PCA[,78:93] == "Unfamiliar", na.rm=TRUE)

DF.PCA$has_phd <- ifelse(DF.PCA$Q15.Demo.HighestDegree == "Doctoral",1,0)

DF.PCA$has_phd <- factor(DF.PCA$has_phd, labels=c("no Doctorate","Doctorate"))

#Humanities <- c("Anthropology", "Arts", "Classics", "History", "Languages", "Literature", "Philosophy", "Religion")

#SocialSciences <- c("Archaeology", "Communication studies", "Cultural and ethnic studies", "Economics", "Geography", "History", "Information science", "Linguistics", "Political science", "Psychology", "Sociology")

#LifeSciences <- c("Biology", "Chemistry", "Earth sciences", "Physics", "Space sciences")

#FormalSciences <- c("Mathematics", "Computer sciences", "Logic", "Statistics", "Systems science")

#Professional <- c("Architecture and design", "Business", "Divinity", "Education", "Engineering", "Human physical performance and recreation", "Journalism, media studies and communication", "Law", "Library and museum studies", "Medicine", "Military sciences", "Public administration")

#Other <- c("Other")

DF.PCA$FieldGroups <- 
  ifelse(DF.PCA$Q5.Demo.PrimaryAcademicField %in% Humanities, 1, 
         ifelse(DF.PCA$Q5.Demo.PrimaryAcademicField %in% SocialSciences, 2,
                ifelse(DF.PCA$Q5.Demo.PrimaryAcademicField %in% LifeSciences, 3,
                       ifelse(DF.PCA$Q5.Demo.PrimaryAcademicField %in% FormalSciences, 4,
                              ifelse(DF.PCA$Q5.Demo.PrimaryAcademicField %in% Professional, 5, 6)))))

DF.PCA$FieldGroups <- factor(DF.PCA$FieldGroups, 
                                levels=c(1:6),
                                labels=c("Humanities", "Social Sciences", "Life Sciences", "Formal Sciences", "Professional", "Other"))

```

```{r}

# Notes on running a factor analysis on categorical data

# suggestions from http://stats.stackexchange.com/questions/5774/can-principal-component-analysis-be-applied-to-datasets-containing-a-mix-of-cont

# from https://en.wikipedia.org/wiki/Multiple_correspondence_analysis

# idea #1: convert current matrix to one where every categorical variable is converted to a series of binary/dummy variables for each possible response, then run Correspondence Analysis

# idea #2: create crosstab table between all of the categorical variables (a "Burt table"), then run Correspondence Analysis

# MCA can also be run using FactoMineR

# http://factominer.free.fr/classical-methods/multiple-correspondence-analysis.html

# Multiple Factor Analysis, however, pays attention to groups of variables. Built into FactoMineR (AFDM())

# http://factominer.free.fr/advanced-methods/multiple-factor-analysis.html
```

## Try using missMDA to impute missing data; won't get used in actually plotting the data?  Then use for additional analyses.

```{r}
# should be able to estimate number of components needed for imputation, but doesn't work

#nb <- estim_ncpMCA(DF.PCA.NA[,c(44:77)])
#result: Error in tab.disj.comp - vrai.tab : non-conformable arrays

complete.MCA <- imputeMCA(DF.PCA.NA[,c(44:77)])
#complete.3 <- imputeMCA(DF.PCA.NA[,c(44:77)],ncp=3)
#complete.4 <- imputeMCA(DF.PCA.NA,ncp=4)
#complete.5 <- imputeMCA(DF.PCA.NA,ncp=5)

res.MCA.imputed <- MCA(DF.PCA.NA, tab.disj = complete.MCA$tab.disj, quali.sup=c(1:43,78:93,104:105), quanti.sup=c(94:103))

#dimdesc(res.MCA.imputed)

#summary(res.MCA.imputed, nbelements = Inf, file="summary_MCA_imputed.txt")

#clust.MCA.imputed <- HCPC(res.MCA.imputed)
clust.MCA.imputed <- HCPC(res.MCA.imputed, nb.clust=4)

clust.MCA.imputed.assign <- clust.MCA.imputed$call$X[,6,drop=FALSE]

names(clust.MCA.imputed.assign) <- "MCA.imputed"

```

```{r}
res.MCA.NA <- MCA(DF.PCA.NA, quali.sup=c(1:43,78:93,104:105),quanti.sup=c(94:103))
                       
#clust.MCA.NA <- HCPC(res.MCA.NA)
clust.MCA.NA <- HCPC(res.MCA.NA, nb.clust=4)
                       
clust.MCA.NA.assign <- clust.MCA.NA$call$X[,6,drop=FALSE]
                       
names(clust.MCA.NA.assign) <- "MCA.NA"
```

```{r}
res.MCA.Skipped <- MCA(DF.PCA, quali.sup=c(1:43,78:93,104:105),quanti.sup=c(94:103)) 

#clust.MCA.Skipped <- HCPC(res.MCA.Skipped)
clust.MCA.Skipped <- HCPC(res.MCA.Skipped, nb.clust=4)


clust.MCA.Skipped.assign <- clust.MCA.Skipped$call$X[,6,drop=FALSE]

names(clust.MCA.Skipped.assign) <- "MCA.Skipped"
```

```{r}

# Note: can't run MFA with both supplemental variables and imputed data

res.MFA.NA <- MFA(DF.PCA.NA, group=c(1,1,2,2,1,18,18,17,17,16,10,2), type=c(rep("n",10),"s","n"), name.group=c("ac.field","high.degree","cons.prod","res.analysis.vis","freq.prod","analysis.tools","vis.tools","estimable.measure","important.measure","layouts","answer.summary","grouped_demo"),num.group.sup=c(1:7,10:12))

#clust.MFA.NA <- HCPC(res.MFA.NA)
clust.MFA.NA <- HCPC(res.MFA.NA, nb.clust=4)

clust.MFA.NA.assign <- clust.MFA.NA$call$X[,6,drop=FALSE]

names(clust.MFA.NA.assign) <- "MFA.NA"
```

```{r}
# impute and cluster just the active variables

complete.MFA.active <- imputeMFA(DF.PCA.NA[,c(44:77)], group=c(17,17), type=(rep("n",2)), name.group=c("estimable.measure","important.measure"))

res.MFA.actimp <- MFA(DF.PCA.NA[,c(44:77)], group=c(17,17), type=(rep("n",2)), name.group=c("estimable.measure","important.measure"),tab.comp = complete.MFA.active)

#clust.MFA.actimp <- HCPC(res.MFA.actimp)
clust.MFA.actimp <- HCPC(res.MFA.actimp, nb.clust=4)

clust.MFA.actimp.assign <- clust.MFA.actimp$call$X[,6,drop=FALSE]

names(clust.MFA.actimp.assign) <- "MFA.actimp"
```

```{r}
# impute and cluster entire table, keeping all variables active; this produces wildly different clustering results

complete.MFA <- imputeMFA(DF.PCA.NA, group=c(1,1,2,2,1,18,18,17,17,16,10,2), type=c(rep("n",10),"s","n"), name.group=c("ac.field","high.degree","cons.prod","res.analysis.vis","freq.prod","analysis.tools","vis.tools","estimable.measure","important.measure","layouts","answer.summary","grouped_demo"))

res.MFA.imputed <- MFA(DF.PCA.NA, group=c(1,1,2,2,1,18,18,17,17,16,10,2), type=c(rep("n",10),"s","n"), name.group=c("ac.field","high.degree","cons.prod","res.analysis.vis","freq.prod","analysis.tools","vis.tools","estimable.measure","important.measure","layouts","answer.summary","grouped_demo"),tab.comp=complete.MFA)

#clust.MFA.imputed <- HCPC(res.MFA.imputed)
clust.MFA.imputed <- HCPC(res.MFA.imputed, nb.clust=4)


clust.MFA.imputed.assign <- clust.MFA.imputed$call$X[,6,drop=FALSE]

names(clust.MFA.imputed.assign) <- "MFA.imputed"
```

```{r}
# run MFA on "Skipped" data

res.MFA.Skipped <- MFA(DF.PCA, group=c(1,1,2,2,1,18,18,17,17,16,10,2), type=c(rep("n",10),"s","n"), name.group=c("ac.field","high.degree","cons.prod","res.analysis.vis","freq.prod","analysis.tools","vis.tools","estimable.measure","important.measure","layouts","answer.summary","grouped_demo"),num.group.sup=c(1:7,10:12))

#clust.MFA.Skipped <- HCPC(res.MFA.Skipped)
clust.MFA.Skipped <- HCPC(res.MFA.Skipped, nb.clust=4)

clust.MFA.Skipped.assign <- clust.MFA.Skipped$call$X[,6,drop=FALSE]

names(clust.MFA.Skipped.assign) <- "MFA.Skipped"
```

```{r}
# different grouping; didn't change clusters

res.MFA.Skipped.group2 <- MFA(DF.PCA, group=c(7,18,18,17,17,16,10,2), type=c(rep("n",6),"s","n"), name.group=c("demo","analysis.tools","vis.tools","estimable.measure","important.measure","layouts","answer.summary","grouped_demo"),num.group.sup=c(1:3,6:8))

#clust.MFA.Skipped.group2 <- HCPC(res.MFA.Skipped.group2)
clust.MFA.Skipped.group2 <- HCPC(res.MFA.Skipped.group2, nb.clust=4)

clust.MFA.Skipped.group2.assign <- clust.MFA.Skipped.group2$call$X[,6,drop=FALSE]

names(clust.MFA.Skipped.group2.assign) <- "MFA.Skipped.group2"
```



```{r}
all.assignments <- merge(clust.MCA.imputed.assign, clust.MCA.NA.assign, by="row.names")

rownames(all.assignments) <- all.assignments[,1]

all.assignments <- all.assignments[,-1]

all.assignments <- merge(all.assignments, clust.MCA.Skipped.assign, by="row.names")

rownames(all.assignments) <- all.assignments[,1]

all.assignments <- all.assignments[,-1]

all.assignments <- merge(all.assignments, clust.MFA.actimp.assign, by="row.names")

rownames(all.assignments) <- all.assignments[,1]

all.assignments <- all.assignments[,-1]

# not using MFA.imputed

#all.assignments <- merge(all.assignments, clust.MFA.imputed.assign, by="row.names")

#rownames(all.assignments) <- all.assignments[,1]

#all.assignments <- all.assignments[,-1]

all.assignments <- merge(all.assignments, clust.MFA.NA.assign, by="row.names")

rownames(all.assignments) <- all.assignments[,1]

all.assignments <- all.assignments[,-1]

all.assignments <- merge(all.assignments, clust.MFA.Skipped.assign, by="row.names")

rownames(all.assignments) <- all.assignments[,1]

all.assignments <- all.assignments[,-1]

all.assignments <- merge(all.assignments, clust.MFA.Skipped.group2.assign, by="row.names")

rownames(all.assignments) <- all.assignments[,1]

all.assignments <- all.assignments[,-1]

for(col in 1:7) {
  all.assignments[,col] <- as.numeric(all.assignments[,col])
}
```

```{r}

all.assignments$agreement <- ifelse(rowMeans(all.assignments[,1:7]) %% 1 > 0.0, FALSE, TRUE)

all.assignments$mode <- apply(all.assignments[,1:7], 1, 
                                   function(idx) { which(tabulate(idx) == max(tabulate(idx)))})



```

```{r}

# group all.assignments by mode (i.e., cluster number) and count participants

cluster.data <- dplyr::rename(all.assignments, cluster=mode) %>% group_by(cluster) %>% summarise(num.participants=n())

```

```{r}

#output final cluster assignments

final_assignments <- dplyr::select(tibble::rownames_to_column(all.assignments), rowname, mode)

final_assignments <- dplyr::rename(final_assignments,V1.Demo.ResponseID=rowname,cluster=mode)

write.csv(final_assignments, file = "final_assignments.csv",row.names=FALSE)

```

```{r}
#dimdesc(res.MFA.NA)

#summary(res.MFA.NA, nbelements = Inf, file="summary_MFA.txt")

#clust.MFA.NA$desc.var
#clust.MFA.NA$desc.ind


# maybe use ggplot instead?
# http://stackoverflow.com/questions/10252639/pca-factominer-plot-data

# extract some parts for plotting
PC1 <- res.MFA.NA$ind$coord[,1]
PC2 <- res.MFA.NA$ind$coord[,2]
PC3 <- res.MFA.NA$ind$coord[,3]
PC4 <- res.MFA.NA$ind$coord[,4]
PC5 <- res.MFA.NA$ind$coord[,5]
labs <- rownames(res.MFA.NA$ind$coord)
PCs <- data.frame(cbind(PC1,PC2,PC3,PC4,PC5))
rownames(PCs) <- labs

data_with_clusts <- merge(DF.PCA.NA, clust.MFA.NA.assign, by="row.names")

rownames(data_with_clusts) <- data_with_clusts[,1]

data_with_clusts <- data_with_clusts[,-1]

data_with_clusts <- merge(data_with_clusts, PCs, by="row.names")

rownames(data_with_clusts) <- data_with_clusts[,1]

data_with_clusts <- data_with_clusts[,-1]

data_with_clusts$extremes <- rowSums(data_with_clusts[,c(96,99)])
```

```{r}

# extract some parts for plotting
var1 <- res.MFA.NA$quanti.var.sup$coord[,1]
var2 <- res.MFA.NA$quanti.var.sup$coord[,2]
var3 <- res.MFA.NA$quanti.var.sup$coord[,3]
var4 <- res.MFA.NA$quanti.var.sup$coord[,4]
var5 <- res.MFA.NA$quanti.var.sup$coord[,5]
vars <- data.frame(cbind(var1,var2,var3,var4,var5))
#vars$names <- rownames(res.MFA.NA$quanti.var.sup$coord)
vars$names <- c("Total Skipped","Measures Skipped","Very High Responses", "Somewhat High Responses","Somewhat Low Responses","Very Low Responses", "Measures Unfamiliar", "Analysis Tools Unfamiliar", "Visualization Tools Unfamiliar", "Layouts Unfamiliar")

ggplot(vars, aes(var1,var2)) + 
  geom_point(size=10,alpha=.7) + 
  geom_text(aes(label=names)) +
  labs(#title="Layout title.",
       x="Dimension 1",
       y="Dimension 2") +
  theme_bw()

```

```{r 'clust-dim12', fig.width=6.5, fig.height=6}
# plots of clustering results compared to demographic variables, etc.

clust_dim12 <- 
ggplot(data_with_clusts) + 
  geom_point(aes(PC1,PC2,col=MFA.NA),size=10,alpha=.7) + 
  scale_color_discrete(name="Cluster") +
  labs(#title="Layout title.",
       x="Dimension 1",
       y="Dimension 2") +
  #scale_x_reverse() +
  theme_bw(base_size = 18)

clust_dim12

ggsave("clust_dim12.png", clust_dim12, width = 10, height = 9)

```

```{r 'biplot-with-arrows', fig.width=6.5, fig.height=6}
# plots of clustering results compared to demographic variables, etc.

biplot_with_arrows <- 
ggplot() + 
  geom_point(data=data_with_clusts, aes(PC1,PC2,col=MFA.NA),size=10,alpha=.7) + 
  #geom_point(data=vars, aes(var1,var2)) +
  geom_segment(data = vars[c(3:6),], aes(xend = var1, yend=var2),
              x=0, y=0, colour="black",
              arrow=arrow(angle=20, type="closed")) +
  geom_text(data=vars[c(3:6),], aes(x=1.2*var1,y=1.2*var2, label=names), size=6) +
  scale_color_discrete(name="Cluster") +
  labs(#title="Layout title.",
       x="Dimension 1",
       y="Dimension 2") +
  #scale_x_reverse() +
  theme_bw(base_size = 18)

biplot_with_arrows

ggsave("biplot_with_arrows.png", biplot_with_arrows, width = 10, height = 9)

```

```{r}
# plots of clustering results compared to demographic variables, etc.

ggplot(tibble::rownames_to_column(data_with_clusts), aes(PC1,PC2)) + 
  geom_point(aes(col=MFA.NA), size=5,alpha=.7) + 
  geom_text(aes(label=rowname, size=(factor(rowname=="R_1IsoEfA2NxRGhHX",levels=c(TRUE,FALSE),ordered=TRUE,labels=c("Yes","No"))))) +
  scale_color_discrete(name="Cluster") +
  scale_size_manual(values=c(3,1)) +
  guides(size="none") +
  labs(#title="Layout title.",
       x="Dimension 1",
       y="Dimension 2") +
  theme_bw()

```

```{r 'biplot_with_sizes', fig.width=6.5}

grid_data <- data_with_clusts %>% dplyr::select(PC1, PC2, ActiveVHigh, extremes, MFA.NA)
grid_data$user <- row.names(grid_data)

grid_data_wide <- tidyr::gather(grid_data, "Indicator", "NumResponses", c(3:4))
grid_data_wide$Indicator <- factor(grid_data_wide$Indicator, labels = c("Very High Responses","Extreme Responses"))
  
ggplot(grid_data_wide) + 
  geom_point(aes(PC1,PC2,size=NumResponses,col=MFA.NA),alpha=.7) + 
  scale_color_discrete(name="Cluster") +
  scale_size_continuous(name="Number of Responses") +
  labs(#title="Layout title.",
       x="Dimension 1",
       y="Dimension 2") +
  #scale_x_reverse() +
  theme_bw(base_size = 13) +
  facet_grid(.~Indicator)
  
```

```{r, fig.width=10, fig.height=8}

ggplot(data_with_clusts) + 
  geom_point(aes(PC1,PC2,size=ActiveVLow,col=MFA.NA),alpha=.7) + 
  scale_color_discrete(name="Cluster") +
  scale_size_continuous(name="No. 'Very Low'\nanswers") +
  labs(#title="Layout title.",
       x="Dimension 1",
       y="Dimension 2") +
  theme_bw() 

```

```{r}

#ggplot(data_with_clusts) + geom_point(aes(PC1,PC2,col=ActiveVLow),size=5,alpha=.7) + theme_bw() + scale_color_gradient(low="grey50",high="grey1")

```

```{r, fig.width=10, fig.height=8}

ggplot(data_with_clusts) + 
  geom_point(aes(PC1,PC2,size=extremes,col=MFA.NA),alpha=.7) + 
  scale_color_discrete(name="Cluster") +
  scale_size_continuous(name="No. 'extreme'\nanswers") +
  labs(#title="Layout title.",
       x="Dimension 1",
       y="Dimension 2") +
  theme_bw()

```

```{r 'scatter-pc1', fig.width=3.25, fig.height=2.5}

modelpc1 <- lm(formula = PC1 ~ sqrt(ActiveVLow), data = data_with_clusts)
r2pc1 <- format(summary(modelpc1)$r.squared, digits = 3)

modelpc1plot <- 
  ggplot(data_with_clusts) + 
  geom_point(aes(y=PC1,x=sqrt(ActiveVLow), fill=MFA.NA), alpha=.7, size=4, color="gray", pch=21) + 
  geom_smooth(aes(x=sqrt(ActiveVLow), y=PC1), method = "lm", se=FALSE, color = "gray") + 
  geom_text(y=3,x=5.2, label=paste("R^2 == ", r2pc1), parse=TRUE, size=4) +
  scale_fill_manual(name="Cluster", values=c("#fef0d9","#fdcc8a","#fc8d59","#d7301f")) +
  #guides(color = "none") +
  scale_y_reverse(limits=c(3.5,-3.5)) + 
  scale_x_continuous(limits=c(0,5.7)) + 
  labs(title="A. Explaining variation in Dimension 1",
       y="Dimension 1",
       x="Square Root of Number of Very Low Responses") +
  theme_bw(base_size = 14)

modelpc1plot

```


```{r 'scatter-pc2', fig.width=3.25, fig.height=2.5}

modelpc2 <- lm(formula = PC2 ~ sqrt(ActiveSLow), data = data_with_clusts)
r2pc2 <- format(summary(modelpc2)$r.squared, digits = 3)

modelpc2plot <- 
  ggplot(data_with_clusts) + 
  geom_point(aes(y=PC2,x=sqrt(ActiveSLow), fill=MFA.NA), alpha=.7, size=4, color="gray", pch=21) + 
  geom_smooth(aes(x=sqrt(ActiveSLow), y=PC2), method = "lm", se=FALSE, color = "gray") + 
  geom_text(x=5.2,y=3, label=paste("R^2 == ", r2pc2), parse=TRUE, size=4) +
  scale_fill_manual(name="Cluster", values=c("#fef0d9","#fdcc8a","#fc8d59","#d7301f")) +
  scale_y_continuous(limits=c(-3.5,3.5)) +
  scale_x_continuous(limits=c(0,5.7)) + 
  labs(title="B. Explaining variation in Dimension 2",
       x="Square Root of Number of Somewhat Low Responses",
       y="Dimension 2") +
  theme_bw(base_size = 14)

modelpc2plot

```

```{r 'grid_pcplot', fig.width=6.5, fig.height=3}

grid.arrange(modelpc1plot, modelpc2plot, ncol=2)

multi_panel_pcplot <- arrangeGrob(modelpc1plot, modelpc2plot, ncol=2)

ggsave("multi_panel_pcplot.png", multi_panel_pcplot, width = 12, height = 5)

```

```{r}

ggplot(data_with_clusts) + 
  geom_boxplot(aes(x=1,y=extremes,col=MFA.NA)) +
  facet_grid(MFA.NA~.) +
  labs(x="Cluster",y="Number of 'extreme' answers") +
  guides(col="none") +
  coord_flip() +
  #scale_x_reverse() + 
  theme_bw() +
  theme(axis.text.y = element_blank(),axis.ticks.y = element_blank(),panel.grid.major.y = element_blank(),panel.grid.minor.y = element_blank())

```

```{r}

ggplot(data_with_clusts) + 
  geom_histogram(aes(extremes,fill=MFA.NA),binwidth=3) +
  facet_grid(MFA.NA~.) +
  guides(fill="none") +
  labs(x="Number of 'extreme' answers (binned)", y="Number of participants") +
  theme_bw()

```

```{r 'responses_by_clust', fig.width=6.5}
responses_by_clust <- 
dplyr::select(data_with_clusts, MFA.NA, Estimable_AvgDegree, Estimable_AvgPathLength, Estimable_AvgShortestPath, Estimable_ClosenessCentrality, Estimable_ClusteringCoeff, Estimable_ComponentSize, Estimable_Density, Estimable_Diameter, Estimable_EigenvectorCentrality, Estimable_LinkBC, Estimable_Loops, Estimable_Modularity, Estimable_NodeBC, Estimable_NodeDegree, Estimable_NumComponents, Estimable_NumLinks, Estimable_NumNodes, Important_AvgDegree, Important_AvgPathLength, Important_AvgShortestPath, Important_ClosenessCentrality, Important_ClusteringCoeff, Important_ComponentSize, Important_Density, Important_Diameter, Important_EigenvectorCentrality, Important_LinkBC, Important_Loops, Important_Modularity, Important_NodeBC, Important_NodeDegree, Important_NumComponents, Important_NumLinks, Important_NumNodes) %>% 
  tidyr::gather("name","answer",2:35) %>% 
  tidyr::separate(name, c("subgroup","name"), sep="_") %>% 
  dplyr::mutate(answer = ifelse(is.na(answer), "Skipped", answer)) %>%
  ggplot(.) + 
  geom_bar(aes(factor(answer,levels=c("Skipped","Unfamiliar","Very low","Somewhat low","Somewhat high","Very high"),ordered=TRUE),fill=MFA.NA)) +
  facet_grid(MFA.NA~.) +
  guides(fill="none") +
  labs(x="Responses to Importance and Estimability Questions", y="Number of Responses") +
  theme_bw(base_size=18)

responses_by_clust

ggsave("responses_by_clust.png", responses_by_clust, width = 10, height = 8)

```

```{r}

# according to res.hcpc$desc.var, HowOftenProduceVis is one of the few demographic variables that correlates strongly with one or more of the clusters.  Try to compare plots of that variable and the cluster assignments?

ggplot(data_with_clusts) + geom_point(aes(PC1,PC2,col=Q11.Demo.HowOftenProduceVis),size=5,alpha=.7) + theme_bw()

```

```{r}

#table(data_with_clusts$Q11.Demo.HowOftenProduceVis,data_with_clusts$MFA.NA)

```

```{r 'prodvis_by_clust', fig.width=6.5}

prodvis_by_clust <-
ggplot(data_with_clusts) + 
  geom_bar(aes(Q11.Demo.HowOftenProduceVis,fill=MFA.NA)) +
  facet_grid(.~MFA.NA) +
  guides(fill="none") +
  coord_flip() + 
  labs(x="How often do you produce network visualizations?", y="Number of Participants") +
  theme_bw(base_size=18)

prodvis_by_clust

ggsave("prodvis_by_clust.png", prodvis_by_clust, width = 10, height = 6)



```

```{r}

#aggregate(data_with_clusts[,112], list(data_with_clusts$MFA.NA), mean)

#plot(res.MFA.NA, invisible=c("var","quali"), select="contrib 3", habillage=7)


```